ANGSD-wrapper: utilities for analyzing next generation sequencing data

### 2016-01-15 manuscript.tex
table, fst

\usepackage[colorinlistoftodos]{todonotes}
\definecolor{flame}{rgb}{0.89, 0.35, 0.13}
%\setlength{\marginparwidth}{1.5cm}
\newcommand{\jri}[1]{\todo[size=\scriptsize, color=flame]{#1}}

\newcommand{\X}{\textcolor{red}{\bf X}}

\title{More than one Author with different Affiliations}
\author[1,$\dag$]{Arun Durvasula}
\author[2,$\dag$]{Paul J. Hoffman}
\author[1]{Tyler V. Kent}
\author[2]{Chaochih Liu}
\author[2]{Thomas J. Y. Kono}
\author[2]{Peter L. Morrell}
\author[1,3,*]{Jeffrey Ross-Ibarra}
\affil[1]{Department of Plant Sciences, University of California, Davis, CA 95616}
\affil[2]{Department of Agronomy and Plant Genetics, University of Minnesota, St. Paul, MN 55108}
\affil[3]{Center for Population Biology and Genome Center, University of California, Davis, CA 95616}
\affil[$\dag$]{These authors contributed equally.}
\affil[*]{email: rossibarra@ucdavis.edu}

@@

\begin{table}
\begin{center}
    \begin{tabular}{ | p{5cm} | p{5cm} |}
    \hline
    \textbf{Methods Implemented} & \textbf{Interactive Graphing} \\ \hline \hline
    Patterson's D (ABBA-BABA) & Yes \\ \hline
    Admixture & Yes  \\ \hline
    Ancestral Sequence & No \\ \hline
    Genotype Likelihoods & No \\ \hline
    Inbreeding Coefficients & No \\ \hline
    PCA & Yes \\ \hline
    Site Frequency Spectrum & Yes \\ \hline
    Diversity Statistics ($\theta_w$, $\theta_\pi$, Fu and Li's $\theta$, Fay's $\theta$)   & Yes \\ \hline
    Neutrality Test Statistics (Tajima's D, Fu and Li's D, Fu and Li's F,  Fay and Wu's H, Zeng's E ) & Yes \\ \hline
    Fst & Yes \\ \hline
    \end{tabular}
    \caption{Table of methods implemented in ANGSD-wrapper}
    \label{tab:methods}
    \end{center}
\end{table}
%\jri{in table: list diversity statistics in parentheses. what about Fst?} done. and added Neutrality test stats

@@

In order to use the plotting software, the user navigates to the desired tab and uploads the appropriate file of results. 
The Shiny server automatically parses ANGSD output files and creates the resulting plot(s) (Figure \ref{fig:theta}), which can be saved using the browser's built in image saving capabilities. % \jri{add genes from gff. i don't think it's Mb 1-10, pls check. if so, is x-axis wrong?} done

@@

\section*{Example data}
% \jri{show k=2 and k=3 for adxmxture in supp.} done
\begin{figure}
\centering
\includegraphics[width=\linewidth]{figures/MainTextWorkflow.pdf}
\caption{Example analysis workflow diagram.}
\label{fig:workflow}
\end{figure}

@@

% \jri{change Mex to mex and make italics in figure axis} done

@@

Consistent with previous results \citep{hufford2012comparative}, we show that the maize SFS is skewed towards more intermediate-frequency variants (Figure \ref{fig:figure3}A-B), likely a result of the bottleneck associated with maize domestication \citep{Beissinger031666}.
We find further evidence of the effect of domestication using the Thetas method, revealing lower levels overall levels of diversity in maize (Figure \ref{fig:figure3}C). These results use Tripsacum as an ancestral sequence, which was generated using the Ancestral Sequence method in ANGSD-wrapper. %\jri{fill in \X's with which A-W script we used. probably need to modify this to mention Tripsacum and generation of the ancestral seq. and which A-W script we used for that.} done
Using the 2DSFS method, which includes an $F_{ST}$ calculation, we then scan our example region for windows with elevated differentiation between maize and teosinte. 
The mean $F_{ST}$ in this region is 0.013, which is an order of magnitude lower than the genome-wide reported value of 0.11 \citep{hufford2012comparative}.   
The \X windows in the 1\% tail of highest $F_{ST}$ values include genes several genes identified as potential targets of selection during domestication in \citep{hufford2012comparative} \jri{see supp tables in Huff. make supp. table of genes in top 1\%. sorry to add, but this is cool \& demonstrates more stuff A-W can do. don't need a graph}

>>>>

\usepackage[colorinlistoftodos]{todonotes}
\definecolor{flame}{rgb}{0.89, 0.35, 0.13}
%\setlength{\marginparwidth}{1.5cm}
\newcommand{\jri}[1]{\todo[size=\scriptsize, color=flame]{#1}}
\newcommand{\fst}{$\textrm{F_{\textrm{\scriptsize ST}}}$}          
\newcommand{\X}{\textcolor{red}{\bf X}}

%\title{More than one Author with different Affiliations}
\author[1,$\dag$]{Arun Durvasula}
\author[2,$\dag$]{Paul J. Hoffman}
\author[1]{Tyler V. Kent}
\author[2]{Chaochih Liu}
\author[2]{Thomas J. Y. Kono}
\author[2]{Peter L. Morrell}
\author[1,3,*]{Jeffrey Ross-Ibarra}
\affil[1]{Department of Plant Sciences, University of California, Davis, CA 95616}
\affil[2]{Department of Agronomy and Plant Genetics, University of Minnesota, St. Paul, MN 55108}
\affil[3]{Center for Population Biology and Genome Center, University of California, Davis, CA 95616}
\affil[$\dag$]{These authors contributed equally.}
\affil[*]{email: rossibarra@ucdavis.edu}

@@

\begin{table}
\begin{center}
    \begin{tabular}{ | p{3.5cm} | p{5cm} | p{2.5cm}  |}
    \hline
    \textbf{Method} & \textbf{Calculations} & \textbf{Interactive Graphing} \\ \hline \hline
    SFS     &    Site frequency spectrum  & Yes  \\ \hline     2DSFS   &    Joint site frequency spectrum, \fst & Yes  \\ \hline 
    Abbababa  &  Patterson's D statistic & Yes  \\ \hline 
    Ancestral &  Extract ancestral sequence from BAM file  & No \\ \hline 
    Genotypes &  Genotype likelihood estimations & No  \\ \hline 
    PCA     &    Principal component analysis & Yes  \\ \hline 
    Thetas   &   Diversity statistics ($\theta_w$, $\theta_\pi$, Fu and Li's $\theta$, Fay's $\theta$) and neutrality tests (Tajima's D, Fu and Li's D, Fu and Li's F,  Fay and Wu's H, Zeng's E ) & Yes  \\ \hline 
    Inbreeding & Calculate per-individual inbreeding coefficients with ngsF & No  \\ \hline 
    Admixture  & Perform admixture analysis & Yes  \\ \hline

 %   Patterson's D (ABBA-BABA) & Yes \\ \hline
%    Admixture & Yes  \\ \hline
%    Ancestral Sequence & No \\ \hline
%    Genotype Likelihoods & No \\ \hline
%    Inbreeding Coefficients & No \\ \hline
%    PCA & Yes \\ \hline
%    Site Frequency Spectrum & Yes \\ \hline
%    Diversity Statistics ($\theta_w$, $\theta_\pi$, Fu and Li's $\theta$, Fay's $\theta$)   & Yes \\ \hline
%    Neutrality Test Statistics (Tajima's D, Fu and Li's D, Fu and Li's F,  Fay and Wu's H, Zeng's E ) & Yes \\ \hline
%    $\textrm{F_{\textrm{\scriptsize ST}}}$ & Yes \\ \hline
    \end{tabular}
    \caption{Table of methods implemented in ANGSD-wrapper}
    \label{tab:methods}
    \end{center}
\end{table}

@@

In order to use the plotting software, the user navigates to the desired tab and uploads the appropriate file of results. 
The Shiny server automatically parses ANGSD output files and creates the resulting plot(s) (Figure \ref{fig:theta}), which can be saved using the browser's built in image saving capabilities.

@@

\section*{Example data}
\begin{figure}
\centering
\includegraphics[width=\linewidth]{figures/MainTextWorkflow.pdf}
\caption{Example analysis workflow diagram.}
\label{fig:workflow}
\end{figure}

@@

@@

Consistent with previous results \citep{hufford2012comparative}, we show that the maize SFS is skewed towards more intermediate-frequency variants (Figure \ref{fig:figure3}A-B), likely a result of the bottleneck associated with maize domestication \citep{Beissinger031666}.
We find further evidence of the effect of domestication using the Thetas method, revealing lower levels overall levels of diversity in maize (Figure \ref{fig:figure3}C). These results use Tripsacum as an ancestral sequence, which was generated using the Ancestral Sequence method in ANGSD-wrapper. %\jri{fill in \X's with which A-W script we used. probably need to modify this to mention Tripsacum and generation of the ancestral seq. and which A-W script we used for that.} done
Using the 2DSFS method, which includes an \fst calculation, we then scan our example region for windows with elevated differentiation between maize and teosinte. 
The mean \fst in this region is 0.013, which is an order of magnitude lower than the genome-wide reported value of 0.11 \citep{hufford2012comparative}.   
The \X windows in the 1\% tail of highest \fst values include genes several genes identified as potential targets of selection during domestication in \citep{hufford2012comparative} \jri{see supp tables in Huff. make supp. table of genes in top 1\%. sorry to add, but this is cool \& demonstrates more stuff A-W can do. don't need a graph}
